Effects of Water Permeability on the Hydrodynamics of a Vesicle Under Flow and Confinement

### 2020-12-28 vesicle_code/docs/permeableVes/main.tex
worked on fig 1 and 7

%
%\paragraph{Literature to maybe cite}
%\cite{ali-ovr-ran2020} show how beading can occur on nanotubes. This
%happens in regions where proteins congregate and this can lead to
%semi-permeability?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
  \centering
   \includegraphics[width=0.475\textwidth]{figures/sketch_sample.png}
  \caption{\label{fig:sketch} Schematics showing the effects of semipermeability on vesicle hydrodynamics in various configurations.}
\end{figure}
\section*{Formulation}
We consider a two-dimensional semi-permeable vesicle (permeable only to water) suspended in a viscous fluid, see figure~\ref{fig:sketch}.
To focus on the effect of water flow on vesicle hydrodynamics in various conditions, we assume the fluid is the same inside and outside the vesicle membrane ($\gamma$) with an outward normal $\nn$.
On the membrane the permeable water flux gives rise to a difference between the fluid velocity ${\bf u}({\bf x}\in \gamma)$
and the membrane velocity $\dot{\bf x}$ \cite{yao-mor2017}:
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - k_w (R T \triangle c + \ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $k_w$ is the water permeability, $R$ is the ideal gas constant, $T$ is the temperature ($^{\circ}$K), $\triangle c$ is contrast in a solute concentration (mol/volume),
and $\ff_\mathrm{mem}$ is the membrane stress that consists of both bending $ \ff_\mathrm{ben}= -k_b \xx_{ssss}$ and tension $\ff_\mathrm{ten}=(\sigma \xx_s)_s$ ($s$
is the arclength along the membrane, see figure~\ref{fig:sketch}. Using $k_w\sim 10^{-12}$ m$^2$s/kg (typical of mammalian cell membranes \cite{LiMoriSun2015_PRL}) and an osmolarity $\triangle c = 0.1$ mol/L (consistent with the osmotic forces in vesicle aspiration experiments \cite{OlbrichRawiczNeedhamEtAl2000_BJ}), the corresponding water flow is of the order $v_\mathrm{o}=0.1 \mu$m/s. Assuming that the membrane bending stiffness $k_b = 10^{-19}$J, a membrane tension $10^{-3}$N/m (one tenth of the lysis tension) and a characteristic length scale of $10^{-6}$m, the tension contribution ($v_\mathrm{t}$) and elastic stress contribution 
($v_\mathrm{e}$) to the permeable water flow can be scaled to $v_\mathrm{o}$ as $v_\mathrm{o}:v_\mathrm{t}:v_\mathrm{e}=1:10^{-2}:10^{-6}$.  Furthermore, we note that in the absence of an osmotic stress, the water permeable flow can be significant when the membrane tension is large due to the external stress such as the hydrodynamics stress due to an external flow and/or due to confinement.
In the following we will explore both.

@@

>>>>

%
%\paragraph{Literature to maybe cite}
%\cite{ali-ovr-ran2020} show how beading can occur on nanotubes. This
%happens in regions where proteins congregate and this can lead to
%semi-permeability?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figures/schematic.pdf}
  \caption{\label{fig:sketch} Schematics showing the effects of
  semipermeability on vesicle hydrodynamics in various configurations.
  The slipper in (a) is the steady-state shape of a semi-permeable
  vesicle in a Poiseuille flow.}
\end{figure}
\section*{Formulation}
We consider a two-dimensional semi-permeable vesicle (permeable only to water) suspended in a viscous fluid, see figure~\ref{fig:sketch}.
To focus on the effect of water flow on vesicle hydrodynamics in various conditions, we assume the fluid is the same inside and outside the vesicle membrane ($\gamma$) with an outward normal $\nn$.
On the membrane the permeable water flux gives rise to a difference between the fluid velocity ${\bf u}({\bf x}\in \gamma)$
and the membrane velocity $\dot{\bf x}$ \cite{yao-mor2017}:
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - k_w (R T \triangle c + \ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $k_w$ is the water permeability, $R$ is the ideal gas constant, $T$ is the temperature ($^{\circ}$K), $\triangle c$ is contrast in a solute concentration (mol/volume),
and $\ff_\mathrm{mem}$ is the membrane stress that consists of both bending $ \ff_\mathrm{ben}= -k_b \xx_{ssss}$ and tension $\ff_\mathrm{ten}=(\sigma \xx_s)_s$ ($s$
is the arclength along the membrane, see figure~\ref{fig:sketch}. Using $k_w\sim 10^{-12}$ m$^2$s/kg (typical of mammalian cell membranes \cite{LiMoriSun2015_PRL}) and an osmolarity $\triangle c = 0.1$ mol/L (consistent with the osmotic forces in vesicle aspiration experiments \cite{OlbrichRawiczNeedhamEtAl2000_BJ}), the corresponding water flow is of the order $v_\mathrm{o}=0.1 \mu$m/s. Assuming that the membrane bending stiffness $k_b = 10^{-19}$J, a membrane tension $10^{-3}$N/m (one tenth of the lysis tension) and a characteristic length scale of $10^{-6}$m, the tension contribution ($v_\mathrm{t}$) and elastic stress contribution 
($v_\mathrm{e}$) to the permeable water flow can be scaled to $v_\mathrm{o}$ as $v_\mathrm{o}:v_\mathrm{t}:v_\mathrm{e}=1:10^{-2}:10^{-6}$.  Furthermore, we note that in the absence of an osmotic stress, the water permeable flow can be significant when the membrane tension is large due to the external stress such as the hydrodynamics stress due to an external flow and/or due to confinement.
In the following we will explore both.

@@

\begin{figure}[htp]
  \ifTikz
  \input{figures/contractingComposite2.tikz}
  \else
  \includegraphics{figures/contractingComposite2.pdf}
  \fi
  \caption{\label{fig:contractingComposite2} Vesicle reduced area and
  velocity vs.~time.}
\end{figure}
